A fast spectral method for the inelastic Boltzmann collision operator and application to heated granular gases\tnoteref{ack}

### 2018-09-10 paper/archive/response/Response to Reviewers.tex

\documentclass[11pt]{article}
\usepackage{latexsym}
\usepackage{amssymb,amsbsy,amsmath,amsfonts,amssymb,amscd}
\usepackage{mathrsfs}
\usepackage{epsfig, graphicx}

\setlength{\oddsidemargin}{0mm} \setlength{\evensidemargin}{0mm}
\setlength{\topmargin}{5mm} \setlength{\textheight}{22cm}
\setlength{\textwidth}{17cm}
\renewcommand{\baselinestretch}{1.0}
\parindent 10pt

\usepackage[margin=1.0in]{geometry}

\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{cases}
\usepackage{xcolor}

\newcommand{\jh}[1]{\textcolor{red}{JH: #1}}
\newcommand{\mz}[1]{\textcolor{blue}{MZ: #1}}

@@

\noindent {We would like to thank the anonymous reviewers for their valuable suggestions and comments, which greatly help us improve the quality and clarity of the manuscript. In the following, {\bf Q} denote the questions or suggestions raised by the reviewers, and {\bf A} refer to our answers. }

\vspace{0.3in}
\centerline{{\large \bf{Response to Reviewer \#1}}}
\bigskip

\begin{itemize}
\item[{\bf Q1}]  {\it Around lines 53 and 54, the authors mentioned that ``the method by Wu et al. only works for 2D Maxwell molecules and 3D hard sphere...". This is absolutely not true as clearly stated in Ref. 16 (see the paragraph after Eq. 22) that the spectral method can handle the Enskog collision term with a general form of collision kernels, which has been demonstrated by the numerical results for hard-sphere, Maxwell, and soft-potential molecules in Fig. 2 therein.}

\item[{\bf A1}] \mz{The ``direct'' or ``standard'' spectral method indeed can handle general form of collision kernels. However, the method by Wu et al., namely the proposed FAST spectral method based on Carleman representation CANNOT work for general collision kernels. We already checked that paper and believe that, only in the case $B(|x|, |y|)$ is a SEPARABLE funtion of $|x|$, $|y|$, the proposed method will work (all the examples shown in that paper are in $B(|x|, |y|) = |x|^a |y|^b$ this form). Of course one can argue that you can use this separable function to APPROXIMATE ``general'' collision kernel, however, we suspect this is good for $B(|x|, |y|)$ being an arbitrary function. On the contrary, our fast method does not have this constrain since we do not reply on Carleman representation and can work for ``TRUE'' general collision kernel without any appoximation of it.}

@@

\item[{\bf Q2}] {\it In lines 60 and 61, evidence on the advantage of `spherical design' over `Gauss quadrature' should be provided.}

\item[{\bf A2}]  \jh{Add a comparsion} \mz{New numercial test is added to illustrate this conclusion, see Table~8.}

@@

\item[{\bf Q3}] {\it G in Eq. 40 is calculated ``exactly," while the gain term is approximated by quadrature. This may violate the mass conservation. This problem may be removed by calculating G using the same quadrature rule as for the gain term.
}

\item[{\bf A3}] \mz{We have explanations for this at line 130 and Section 4. The conservation can be achieved using other approachn (e.g., adding a correction step).}

@@

\item[{\bf Q4}] {\it In Fig. 2 when comparing the direct method to the Fourier spectral method, the software, computer configurations, number of CPU cores used, etc. should be specified.  }

\item[{\bf A4}] \mz{Added accordingly, see Fig. 2}

\item[{\bf Q5}] {\it The accuracy of the proposed Fourier spectral method is proved in spatially homogeneous relaxation problems. The paper is a little bit short. I think at least one spatially-inhomogeneous problem should be simulated to test the stability of the present method because some deterministic numerical method fails to solve this problem although they can solve spatially homogeneous relaxation problems. }

\item[{\bf A5}] \mz{This is no reason that this method will fail for spatially in-homogeneous problem......}

@@

First of all, this reviewer's comment ``This method is almost the same than the one introduced in the manuscript" is absolutely not true! The method in [16] is based on the so-called Carleman representation of the collision operator. This representation is first used in Mouhot and Pareschi (2006) to derive a fast Fourier method for the classical (elastic) Boltzmann collision operator and is in fact the basis of the method proposed in [16]. Our method is based on the $\sigma$-representation of the collision operator. The two representations are equivalent at the continuous level. However, they become quite different when numerical approximation is applied, for example, the truncation of the integral is done on different variables, in our method, it is $g=v-v_*$, the relative velocity is truncated, while in [16], it is $x=(|g|\sigma-g)/2$ and $y=-g-x$ are truncated; furthermore, our numerical quadrature is applied on $g$, while in [16], it is $x$ that is discretized. These will of course result in different approximations of the collision operator. Another difference is that, our method makes no assumptions on the collision kernel and applies to the general form $B_{\sigma}(|g|,\sigma\cdot \hat{g})$. The method in [16], although can treat general kernels, is based on a modification of the kernel so that it is in the separable form. In view of this, our method does not need any special treatment/approximation of the kernel and is much easier to implement. Moreover, we propose to use the sophisticated quadrature, spherical design, on the sphere, in contrast to the tensor product based Gauss quadrature in [16], which results in much better accuracy as demonstrated in our paper.

\jh{Our method provides better accuracy compared to that in [16], do the same numerical test in Section 4.1 of [16] ($10^{-8}$ v.s. $10^{-5}$). I think [16] does not contain a CPU time, please double check (\mz{Confirmed, at least not comparable}).}

@@

\item[{\bf Q6}] {\it The reviewer doesn't understand AT ALL the assertion at the end of p. 11 saying that the ``integrand in (39) is oscillatory on the scale of $O(N)$". What does that means? More details (a lot !) are needed here!}

\item[{\bf A6}]

\item[{\bf Q7}] {\it On p. 12, expressions for the weight $w_{\rho}$ and $w_g$ are missing, as well as more details on the underlying numerical methods used here!}

\item[{\bf A7}] \mz{They are just standard Gauss-Legendre quardrature weights which are included in any numerical textbook.}

@@

\item[{\bf Q9}] {\it What about the conservations properties of the new scheme? Are there any numerical simulations available?
}

\item[{\bf A9}] The Fourier spectral method does not have the conservation. We did not check the mass and momentum but the extensive results presented regarding the temperature demonstrate that the method can capture the moments fairly accurately.

\item[{\bf Q10}] {\it Have the authors tried to take $e=1$ namely the elastic case, in order to see if the BKW solution is obtained?
}

\item[{\bf A10}] \jh{Maybe include it in Section 4.1.1 as this is quite easy \mz{Of course we checked our code can recover the elastic case otherwise it means our code is wrong....}.}

@@

\item[{\bf Q11}] {\it On p. 14, using spectral to compute the laplacian is nice, but what is the effect on the aliasing properties? Should one needs to change the truncation? this should be studied too, because the Laplace operator spreads in velocity, whereas Q concentrates, and the analysis on R and S is then broken.

}

\item[{\bf A11}] \mz{This is a good question. But since it is just a heating bath with a very small $\varepsilon = 10^{-6}$ in our test, we believe, at least numerical results show that we do not need to change the truncation.}

@@

\item[{\bf Q14}] {\it On p. 6, it seems to the reviewer that it is usually easier to derive the strong form of Q, start- ing from its weak form, rather than the contrary (due to the irreversible microscopic collision dynamics).}

\item[{\bf A14}] This is more of a personal taste. The review paper [4] by Villani also starts with the strong form and then introduces the weak form. In fact, it is quite intuitive to understand the strong form directly: the pre-collisional velocities $\tilde{v}$ and $\tilde{v}_*$ become $v$ and $v_*$ after the collision, hence yields the gain term of the collision operator.

\item[{\bf Q15}] {\it Why does the author decide in the numerical section to use the BKW initial conditions? This has no physical relevancy in the inelastic case (and too much symmetries).}

\item[{\bf A15}] We considered the BKW solution in Section 4.1.1 as a starting example mainly because it is an isotropic solution and becomes an exact solution when $e=1$. Later in Section 4.1.2 we did consider an anisotropic solution (see equation (50)) and observed quite different behavior. \mz{We have added an anisotropic solution case in 3D, see the new subsection~4.2.2.}\jh{Why didn't we consider an anisotropic solution in 3D?}

### 2018-09-14 paper/paper_rev/fast_spec_inelastic_rev.tex
Revised version submitted.

{\bf Related work.} A closely related work is the fast Fourier spectral method introduced in \cite{WZR15} for the generalized Enskog equation (the Boltzmann and Enskog collision operators reduce to the same thing in the spatially homogeneous case). This method is based on a different representation of the collision operator, the so-called Carleman form, and relies on an extra approximation of the integral in the radial direction. As a result, the final computational cost is comparable to our method. However, we point out that the method in \cite{WZR15} only readily works for 2D Maxwell molecules and 3D hard spheres. To deal with other types of interactions (even 3D Maxwell molecules, say), additional modification of the kernel or parameter fitting is necessary, as was done in \cite{WWSRZ13}. In contrast, our method makes no assumption in the collision kernel and is very easy to implement. Further, we propose to use the spherical design \cite{Womersley} to approximate the integral on the unit sphere, which is the optimal quadrature on the sphere up to date and shows better performance (require much fewer points to attain the same level of accuracy) than the tensor product Gauss quadrature as used in \cite{WZR15}.

@@

Now assuming $f$ is a solution to equation \eqref{inBoltz}, the conservation of mass and momentum is straightforward by taking $\varphi = 1$ and $v$ respectively in the weak form \eqref{weak22}, i.e.,
\begin{equation}
\rho \equiv \rho_0, \quad  m\equiv m_0,
\end{equation}
with $\rho_0$ and $m_0$ being the initial density and momentum. By taking $\varphi=|v|^2/2$ in \eqref{weak22}, one obtains the following energy equation
\begin{equation}\label{ki_energy}
\partial_t E -\varepsilon d\rho_0=-\frac{1}{16}\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\int_{S^{d-1}} (1-e^2)B_{\sigma}(|g|,\sigma\cdot \hat{g})|g|^2(1-\sigma\cdot \hat{g})ff_*\,\rd{\sigma}\rd{v}\rd{v_*},
\end{equation}
where we used (\ref{momentum}) and (\ref{relation}).

For constant restitution coefficient $e$ and Maxwell molecules ($\lambda=0$ in (\ref{VHS})), equation \eqref{ki_energy} becomes
\begin{equation} 
\partial_t E -\varepsilon d\rho_0=-\frac{1-e^2}{16}C_0\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\int_{S^{d-1}}  |g|^{2}(1-\sigma\cdot \hat{g})ff_*\,\rd{\sigma}\rd{v}\rd{v_*}.
\end{equation}
Note that
\begin{equation}
\int_{S^{d-1}} C_0(1-\sigma\cdot \hat{g})\,\rd{\sigma}
\end{equation}
is a positive constant regardless of $\hat{g}$ and, without loss of generality, is assumed to be $1$. Then
\begin{align} 
  \partial_t E -\varepsilon d\rho_0 &=-\frac{1-e^2}{16}\int_{\mathbb{R}^d}\int_{\mathbb{R}^d} |g|^{2}ff_*\,\rd{v}\rd{v_*} \notag\\
   & =-\frac{1-e^2}{16}\int_{\mathbb{R}^d}\int_{\mathbb{R}^d} (|v|^2+|v_*|^2-2v\cdot v_*)ff_*\,\rd{v}\rd{v_*} \notag \\
   &=-\frac{1-e^2}{8}\left(2\rho_0E - m_0^2\right).
\end{align}

@@

The starting point is the weak form (\ref{weak11}) with a change of integration variable $v_* \rightarrow g=v-v_*$:
\begin{align} 
\int_{\mathbb{R}^d} Q(f,f)(v)\,\varphi(v)\,\rd{v}&= \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{S^{d-1}} B_{\sigma}(|g|,\sigma\cdot \hat{g})f(v)f(v-g)  \left(\varphi(v')-\varphi(v)\right)\,\rd{\sigma} \rd{g} \rd{v},
\end{align}
where
\begin{equation}
v'=v-\frac{1+e}{4}(g-|g|\sigma).
\end{equation}
It is then assumed that $f$ has a compact support: $\text{Supp}_v(f) \approx B_S$, where $B_S$ is a ball centered at the origin with radius $S$ (in practice, the support can be chosen as some multiple of the initial thermal speed $v_{\text{th}}=\sqrt{T_0}$). Hence it suffices to truncate the infinite integral in $g$ to a larger ball $B_R$ with radius $R=2S$:
\begin{align}  \label{weak}
\int_{\mathbb{R}^d} Q(f,f)(v)\,\varphi(v)\,\rd{v}&= \int_{\mathbb{R}^d} \int_{B_R} \int_{S^{d-1}} B_{\sigma}(|g|,\sigma\cdot \hat{g})f(v)f(v-g)  \left(\varphi(v')-\varphi(v)\right)\,\rd{\sigma} \rd{g} \rd{v}.
\end{align}
Next, we restrict $v$ to a cubic computational domain $D_L=[-L,L]^d$, and approximate $f$ by a truncated Fourier series:
\begin{equation}
f(v)\approx f_N(v)=\sum_{k=-\frac{N}{2}}^{\frac{N}{2}-1}\hat{f}_k e^{i\frac{\pi}{L}k\cdot v}, \quad \hat{f}_k=\frac{1}{(2L)^d}\int_{D_L}f(v)e^{-i\frac{\pi}{L}k\cdot v}\,\rd{v}.
\end{equation}
Note that $k=(k_1,\dots,k_d)$ is a multidimensional index, and $\sum_{k=-N/2}^{N/2-1}:=\sum_{k_1,\dots,k_d=-N/2}^{N/2-1}$. The choice of $L$ can be made similarly as in the classical case $L\geq (3+\sqrt{2})S/2$, see \cite{PR00} for more details. Now substituting $f_N$ into (\ref{weak}) and choosing $\varphi(v)=e^{-i\frac{\pi}{L}k\cdot v}$, we can obtain the $k$-th mode of the collision operator as
\begin{align} \label{QQ}
\hat{Q}_k:&=\frac{1}{(2L)^d}\int_{D_L}Q(f_N,f_N)(v)e^{-i\frac{\pi}{L}k\cdot v}\,\rd{v}\nonumber\\
&=\frac{1}{(2L)^d}\int_{D_L} \int_{B_R} \int_{S^{d-1}} B_{\sigma}(|g|,\sigma\cdot \hat{g})f_N(v)f_N(v-g)  \left(e^{-i\frac{\pi}{L}k\cdot v'}-e^{-i\frac{\pi}{L}k\cdot v}\right)\,\rd{\sigma} \rd{g} \rd{v}\nonumber\\
&=\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\int_{B_R}\int_{S^{d-1}}B_{\sigma}(|g|,\sigma\cdot \hat{g})e^{-i\frac{\pi}{L}m\cdot g}\left(e^{i\frac{\pi}{L}\frac{1+e}{4}k\cdot (g-|g|\sigma)}-1\right)\hat{f}_l\hat{f}_m\,\rd{\sigma} \rd{g},
\end{align}
where in the last equality we used the orthogonality of the Fourier basis.

@@

We choose the same set of parameters as in the previous example and perform the convergence test in $N$. The results are shown in Table~\ref{table 2}. Overall, the accuracy degrades a bit compared to the isotropic solution. Surprisingly, the accuracy of the ``separate" method is much better than the ``full" method when $N$ gets bigger. Although it seems more natural to approximate the gain and loss terms together using the weak formulation, these results suggest that one should use the more accurate formula for the loss term when it is available (intuitively, the smoothing property of the gain term \cite{Lions94} alleviates the error introduced by the quadrature; since the loss term does not have this property, the same quadrature can incur more errors, hence a more accurate formula should be used).

@@

Numerical examples in this subsection are generated using the ``separate" method (the loss term is computed as in (\ref{separate3}).

\subsubsection{3D inelastic Maxwell molecules}

@@

\begin{table}[htp!]
  \centering
  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
    \toprule
    $M_{\text{sph}}$ & $|T_\text{num} - T_\text{ref}|$\\
    \midrule
    6 & 0.0013957739849754791 \\
    12 & 9.9706271716293315e-05 \\
    32 & 2.2499901350947482e-06 \\
    48 & 2.4272557155313734e-06 \\
    70 & 2.4703481364962698e-06 \\
    94 & 2.4703481364962698e-06 \\
    120 & 2.453380453903975e-06 \\
    \bottomrule
  \end{tabu}
  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule for $e=0.2$. Errors shown here are $|T_\text{num} - T_\text{ref}|$ at $t_\text{final}=1$. $\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
  \label{table 3}
\end{table}

\begin{table}[htp!]
  \centering
  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
    \toprule
    $M_{\text{sph}}$ & $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$\\
    \midrule
 	6 & 0.00041820767783433107 \\
	12 & 3.1726851245650724e-05 \\
	32 & 6.5752814213618921e-07 \\
	48 & 5.6132233266191489e-07 \\
	70 & 2.6714628356683257e-07 \\
	94 & 1.0508293902503248e-07 \\
	120 & 2.8872928147741922e-08 \\ 
    \bottomrule
  \end{tabu}
  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule for $e=0.2$. Errors shown here are $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$ at $t_\text{final}=1$. $\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
  \label{table 4}
\end{table}

@@

\begin{table}[H]
	\centering
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$ & direct & fast \\
    \midrule
    8 & 32.3ms & 17.6ms \\
    16 & 602ms & 140ms \\
    32 & 23.7s & 1.18s \\
    64 & -- & 9.7s \\
    \bottomrule
  \end{tabu}
  	\caption{Average running time per evaluation of the collision operator. Comparison between the direct method and the fast method for various $N$ and fixed $N_{\rho}=30$, $M_{\text{sph}}=32$.}
	\label{table 5}
 \end{table}

\begin{figure}[htp!]
  \centering
  \includegraphics[width = .8\linewidth]{CPU_time}
  \caption{Same numbers as in Table~\ref{table 5} plotted in a log-log scale.}
  \label{CPUtime}
\end{figure}

>>>>

{\bf Related work.} A closely related work is the fast Fourier spectral method introduced in \cite{WZR15} for the generalized Enskog equation (the Boltzmann and Enskog collision operators reduce to the same thing in the spatially homogeneous case). This method is based on a different representation of the collision operator, the so-called Carleman form, and relies on an extra approximation of the integral in the radial direction. As a result, the final computational cost is comparable to our method. However, we point out that the method in \cite{WZR15} only readily works for 2D Maxwell molecules and 3D hard spheres \textcolor{red}{as in these two cases the collision kernel is in a separable form so that the resulting sum is a convolution and the fast Fourier transform can be applied (see \cite{MP06} where the Carleman form was first used in the numerical approximation of the Boltzmann collision operator)}. To deal with other types of interactions (even 3D Maxwell molecules, say), additional modification of the kernel or parameter fitting is necessary, as was done in \cite{WWSRZ13}. In contrast, our method makes no assumption in the collision kernel and is very easy to implement. Further, we propose to use the spherical design \cite{Womersley} to approximate the integral on the unit sphere, which is the optimal quadrature on the sphere up to date and shows better performance (require much fewer points to attain the same level of accuracy) than the tensor product Gauss quadrature as used in \cite{WZR15}.

@@

Now assuming $f$ is a solution to equation \eqref{inBoltz}, the conservation of mass and momentum is straightforward by taking $\varphi = 1$ and $v$ respectively in the weak form \eqref{weak22}, i.e.,
\begin{equation}\label{conservation}
\rho \equiv \rho_0, \quad  m\equiv m_0,
\end{equation}
with $\rho_0$ and $m_0$ being the initial density and momentum. By taking $\varphi=|v|^2/2$ in \eqref{weak22}, one obtains the following energy equation
\begin{equation}\label{ki_energy}
\partial_t E -\varepsilon d\rho_0=-\frac{1}{16}\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\int_{S^{d-1}} (1-e^2)B_{\sigma}(|g|,\sigma\cdot \hat{g})|g|^2(1-\sigma\cdot \hat{g})ff_*\,\rd{\sigma}\rd{v}\rd{v_*},
\end{equation}
where we used (\ref{momentum}) and (\ref{relation}).

For constant restitution coefficient $e$ and Maxwell molecules ($\lambda=0$ in (\ref{VHS})), equation \eqref{ki_energy} becomes
\begin{equation} 
\partial_t E -\varepsilon d\rho_0=-\frac{1-e^2}{16}C_0\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\int_{S^{d-1}}  |g|^{2}(1-\sigma\cdot \hat{g})ff_*\,\rd{\sigma}\rd{v}\rd{v_*}.
\end{equation}
Note that
\begin{equation}
\int_{S^{d-1}} C_0(1-\sigma\cdot \hat{g})\,\rd{\sigma}
\end{equation}
is a positive constant regardless of $\hat{g}$ and, without loss of generality, is assumed to be~$1$. Then
\begin{align} 
  \partial_t E -\varepsilon d\rho_0 &=-\frac{1-e^2}{16}\int_{\mathbb{R}^d}\int_{\mathbb{R}^d} |g|^{2}ff_*\,\rd{v}\rd{v_*} \notag\\
   & =-\frac{1-e^2}{16}\int_{\mathbb{R}^d}\int_{\mathbb{R}^d} (|v|^2+|v_*|^2-2v\cdot v_*)ff_*\,\rd{v}\rd{v_*} \notag \\
   &=-\frac{1-e^2}{8}\left(2\rho_0E - m_0^2\right).
\end{align}

@@

The starting point is the weak form (\ref{weak11}) with a change of integration variable $v_* \rightarrow g=v-v_*$:
\begin{align} 
\int_{\mathbb{R}^d} Q(f,f)(v)\,\varphi(v)\,\rd{v}&= \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{S^{d-1}} B_{\sigma}(|g|,\sigma\cdot \hat{g})f(v)f(v-g)  \left(\varphi(v')-\varphi(v)\right)\,\rd{\sigma} \rd{g} \rd{v},
\end{align}
where
\begin{equation}
v'=v-\frac{1+e}{4}(g-|g|\sigma).
\end{equation}
It is then assumed that $f$ has a compact support: $\text{Supp}_v(f) \approx B_S$, where $B_S$ is a ball centered at the origin with radius $S$ (in practice, $S$ can be chosen roughly as  $\max(u_0\pm c\sqrt{T_0})$, where $c$ is some constant around 3). Hence it suffices to truncate the infinite integral in $g$ to a larger ball $B_R$ with radius $R=2S$:
\begin{align}  \label{weak}
\int_{\mathbb{R}^d} Q(f,f)(v)\,\varphi(v)\,\rd{v}&= \int_{\mathbb{R}^d} \int_{B_R} \int_{S^{d-1}} B_{\sigma}(|g|,\sigma\cdot \hat{g})f(v)f(v-g)  \left(\varphi(v')-\varphi(v)\right)\,\rd{\sigma} \rd{g} \rd{v}.
\end{align}
Next, we restrict $v$ to a cubic computational domain $D_L=[-L,L]^d$, and approximate $f$ by a truncated Fourier series:
\begin{equation}
f(v)\approx f_N(v)=\sum_{k=-\frac{N}{2}}^{\frac{N}{2}-1}\hat{f}_k e^{i\frac{\pi}{L}k\cdot v}, \quad \hat{f}_k=\frac{1}{(2L)^d}\int_{D_L}f(v)e^{-i\frac{\pi}{L}k\cdot v}\,\rd{v}.
\end{equation}
Note that $k=(k_1,\dots,k_d)$ is a multidimensional index, and $\sum_{k=-N/2}^{N/2-1}:=\sum_{k_1,\dots,k_d=-N/2}^{N/2-1}$. The choice of $L$ can be made similarly as in the classical case $L\geq (3+\sqrt{2})S/2$, see \cite{PR00, FPT05} for more details. Now substituting $f_N$ into (\ref{weak}) and choosing $\varphi(v)=e^{-i\frac{\pi}{L}k\cdot v}$, we can obtain the $k$-th mode of the collision operator as
\begin{align} \label{QQ}
\hat{Q}_k:&=\frac{1}{(2L)^d}\int_{D_L}Q(f_N,f_N)(v)e^{-i\frac{\pi}{L}k\cdot v}\,\rd{v}\nonumber\\
&=\frac{1}{(2L)^d}\int_{D_L} \int_{B_R} \int_{S^{d-1}} B_{\sigma}(|g|,\sigma\cdot \hat{g})f_N(v)f_N(v-g)  \left(e^{-i\frac{\pi}{L}k\cdot v'}-e^{-i\frac{\pi}{L}k\cdot v}\right)\,\rd{\sigma} \rd{g} \rd{v}\nonumber\\
&=\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\int_{B_R}\int_{S^{d-1}}B_{\sigma}(|g|,\sigma\cdot \hat{g})e^{-i\frac{\pi}{L}m\cdot g}\left(e^{i\frac{\pi}{L}\frac{1+e}{4}k\cdot (g-|g|\sigma)}-1\right)\hat{f}_l\hat{f}_m\,\rd{\sigma} \rd{g},
\end{align}
where in the last equality we used the orthogonality of the Fourier basis.

@@

We choose the same set of parameters as in the previous example and perform the convergence test in $N$. The results are shown in Table~\ref{table 2}. Overall, the accuracy degrades a bit compared to the isotropic solution. Surprisingly, the accuracy of the ``separate" method is much better than the ``full" method when $N$ gets bigger. Although it seems more natural to approximate the gain and loss terms together using the weak formulation, these results suggest that one should use the more accurate formula for the loss term when it is available. Intuitively, the smoothing property of the gain term \textcolor{red}{(see \cite{Lions94} for the classical collision operator and \cite{MM06} for the inelastic case)} alleviates the error introduced by the quadrature; since the loss term does not have this property, the same quadrature can incur more errors, hence a more accurate formula should be used.

@@

Numerical examples in this subsection are generated using the ``separate" method (the loss term is computed as in equation (\ref{separate3})).

\subsubsection{3D inelastic Maxwell molecules \color{red} --- isotropic solution}

@@

\begin{table}[htp!]
  \centering
  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
    \toprule
    $M_{\text{sph}}$ & $|T_\text{num} - T_\text{ref}|$\\
    \midrule
    6 & 0.0013957739849754791 \\
    12 & 9.9706271716293315e-05 \\
    32 & 2.2499901350947482e-06 \\
    48 & 2.4272557155313734e-06 \\
    70 & 2.4703481364962698e-06 \\
    94 & 2.4703481364962698e-06 \\
    120 & 2.453380453903975e-06 \\
    \bottomrule
  \end{tabu}
  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule {\color{red} (isotropic solution)} for $e=0.2$. Errors shown here are $|T_\text{num} - T_\text{ref}|$ at $t_\text{final}=1$. $\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
  \label{table 3}
\end{table}

\begin{table}[htp!]
  \centering
  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
    \toprule
    $M_{\text{sph}}$ & $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$\\
    \midrule
 	  6 & 0.00041820767783433107 \\
	  12 & 3.1726851245650724e-05 \\
	  32 & 6.5752814213618921e-07 \\
	  48 & 5.6132233266191489e-07 \\
	  70 & 2.6714628356683257e-07 \\
	  94 & 1.0508293902503248e-07 \\
	  120 & 2.8872928147741922e-08 \\ 
    \bottomrule
  \end{tabu}
  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule {\color{red} (isotropic solution)} for $e=0.2$. Errors shown here are $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$ at $t_\text{final}=1$. $\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
  \label{table 4}
\end{table}

@@

\begin{table}[H]
	\centering
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$ & direct & fast \\
    \midrule
    8 & 32.3ms & 17.6ms \\
    16 & 602ms & 140ms \\
    32 & 23.7s & 1.18s \\
    64 & -- & 9.7s \\
    \bottomrule
  \end{tabu}
  	\caption{Average running time per evaluation of the collision operator. Comparison between the direct method and the fast method for various $N$ and fixed $N_{\rho}=30$, $M_{\text{sph}}=32$. {\color{blue} This test is performed using Python (numpy and scipy modules for array operations; pyfftw module (a Python wrapper of FFTW) for computing FFTs) on a $10$ CPUs workstation (Intel(R) Xeon(R) CPU E5-2660 v3 @ 2.60GHz, $8$ threads per FFTW operation).}}
	\label{table 5}
 \end{table}

\begin{figure}[htp!]
  \centering
  \includegraphics[width = .8\linewidth]{CPU_time}
  \caption{Same numbers as in Table~\ref{table 5} plotted in a log-log scale. }
  \label{CPUtime}
\end{figure}

{\color{red} \subsubsection{3D inelastic Maxwell molecules  --- anisotropic solution}

In previous subsection we performed a convergence test in $M_\text{sph}$ (compared with the direct method and the analytical solution of the temperature) by considering an isotropic solution. This test shows that by using spherical design one can achieve very high accuracy to approximate the integral on the unit sphere, which is a key step in our 3D algorithm. In order to further investigate the performance of spherical design and to illustrate its necessity in the algorithm, an example with anisotropic solution is presented in this subsection.

For the sake of comparing with isotropic case, this test is done using exactly the same configuration as in previous subsection except the initial condition:
\begin{equation}\label{3D_anisotropic}
  f_0(v) = \frac{4}{5\pi^{\frac{3}{2}}}\left(\exp(-\sqrt[3]{16}|v-2|^2) + \exp(-|v+0.5|^2)\right).
\end{equation}
The results are given in Table~\ref{3D_T_conv_anisotropic} and Table~\ref{3D_direct_conv_anisotropic}. The error decays with increasing $M_{\text{sph}}$ as expected. However, it is also evident that the accuracy is not as good as in the isotropic case. Despite this result, we consider it as acceptable since the problem itself (numerical approximation of an integral on 3D unit sphere) is highly nontrivial and is an active research field. We believe spherical design is so far among the best methods to deal with such problems.} \textcolor{blue}{We illustrate this statement by performing another test using the standard tensor product based Gauss quadrature, which is also the method used in~\cite{WZR15} for the computation of the integral on 3D unit sphere. The results are shown in Table~\ref{3D_quadrature_conv_anisotropic}. Under roughly the same total number of quadrature points, this method has very bad performance: slow convergence and low accuracy.}

\begin{table}[htpb]
  \color{red}
  \centering
  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
    \toprule
    $M_{\text{sph}}$ & $|T_\text{num} - T_\text{ref}|$\\
    \midrule
    6 & 2.661413377541729e-01 \\
    12 & 1.6773023124809094e-01 \\
    32 & 1.0297536867238799e-01 \\
    48 & 2.584202650005074e-02 \\
    70 & 2.422376071694199e-02 \\
    94 & 3.058799838901116e-03 \\
    120 & 4.413127388460447e-03 \\
    156 & 1.220110856786194e-03 \\
    192 & 3.96517639616345e-04 \\
    \bottomrule
  \end{tabu}
  \captionsetup{labelfont={color=red}}
  \caption{\color{red} Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule (anisotropic solution) for $e=0.2$. Errors shown here are $|T_\text{num} - T_\text{ref}|$ at $t_\text{final}=1$. $\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
  \label{3D_T_conv_anisotropic}
\end{table}

\begin{table}[htpb]
  \color{red}
  \centering
  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
    \toprule
    $M_{\text{sph}}$ & $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$\\
    \midrule
 	  6 & 1.4674566108660258e-02 \\
    12 & 9.532325835757156e-03 \\
    32 & 5.92707599850977e-03 \\
    48 & 1.952601851310574e-03 \\
    70 & 1.43696494086e-03 \\
    94 & 4.354500695785647e-04 \\
    120 & 3.534377201557734e-04 \\
    156 & 9.811152052724729e-05 \\
    192 & 3.747408101774895e-05 \\ 
    \bottomrule
  \end{tabu}
  \captionsetup{labelfont={color=red}}
  \caption{\color{red} Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule (anisotropic solution) for $e=0.2$. Errors shown here are $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$ at $t_\text{final}=1$. $\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
  \label{3D_direct_conv_anisotropic}
\end{table}

\begin{table}[htpb]
  \color{blue}
  \centering
  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
    \toprule
    \# of $(\theta, \phi)$ & $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$\\
    \midrule
    (2,3) & 9.301850859331023e-03 \\
    (3,4) & 1.4164542178285952e-02 \\
    (4,8) & 2.148725985732372e-02 \\
    (6,8) & 3.952940827407407e-03 \\
    (7,10) & 3.3540095954422665e-03 \\
    (10,10) & 3.405850043869026e-03 \\
    (10,12) & 3.016912442998702e-03 \\
    (11,15) & 1.227818352872178e-03 \\
    (12,16) & 1.2310868626689182e-03 \\
    \bottomrule
  \end{tabu}
  \captionsetup{labelfont={color=blue}}
  \caption{\color{blue} Convergence test of Gauss quadrature of the 3D Maxwell molecule (anisotropic solution) for $e=0.2$. Errors shown here are $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$ at $t_\text{final}=1$. $\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
  \label{3D_quadrature_conv_anisotropic}
\end{table}
